Plug-and-play dual-tree algorithm runtime analysis

### 2015-10-07 bounds-files/imbalance.tex
Update imbalance numbers with better experiment. That sure did take a long time to run...

However, for most real-world datasets with the cover tree implementation in {\bf
mlpack} \citep{mlpack2013} and the reference implementation
\citep{langford2006}, the tree imbalance is near-linear with the number of
points.  We have constructed cover trees on $N$ uniformly subsampled points
from a variety of datasets and calculated the imbalance; see Table
\ref{tab:imbalance} for the results.  These results are normalized with respect
to $N$, for which the values of $5000$, $50000$, and $500000$ were chosen.
The `power', `susy', `higgs', and `covertype' datasets are found in the UCI
Machine Learning Repository \cite{ucimlrepository}, the `mnist' dataset is from
\cite{lecun2000mnist}, and the `lcdm` and `sdss' datasets are Sloan Digital Sky
Survey data \cite{adelman2008sixth}.  The imbalances on each of these datasets
tend to be near-linear.

@@

\begin{table}[tb]
\begin{center}
\begin{tabular}{| c | c | c | c | c |}
\hline
 & & \multicolumn{3}{|c|}{Imbalance} \\
\hline
Dataset & $d$ & $N = 5k$ & $N = 50k$ & $N = 500k$ \\
\hline
%lcdm & 3 & 29402 & 313190 & 3400305 \\
lcdm & 3 & 5.88 & 6.26 & 6.80 \\
%sdss & 4 & 17691 & 191716 & 1483711 \\
sdss & 4 & 3.54 & 3.83 & 2.97 \\
% power & 7 & 22152 & 202050 & 1345825 \\
power & 7 & 4.43 & 4.04 & 2.69 \\
% susy & 18 & 3686 & 64231 & 930969 \\
susy & 18 & 0.74 & 1.28 & 1.86 \\
% higgs & 29 & 404 & 1272 & 73793 \\
higgs & 29 & 0.08 & 0.03 & 0.15 \\
% covertype & 55 & 9680 & 100344 & 1319447 \\
covertype & 55 & 1.94 & 2.01 & 2.64 \\
% mnist & 784 & 4004 & 79375 & 983554 \\
mnist & 784 & 0.80 & 1.59 & 1.97 \\
\hline
\end{tabular}
\end{center}
\caption{Empirically calculated tree imbalances, normalized by $N$.}
\label{tab:imbalance}
\end{table}

>>>>

However, for most real-world datasets with the cover tree implementation in {\bf
mlpack} \citep{mlpack2013} and the reference implementation
\citep{langford2006}, the tree imbalance is near-linear with the number of
points.  We have constructed cover trees on $N$ uniformly subsampled points
from a variety of datasets and calculated the imbalance; see Table
\ref{tab:imbalance} for the results.  Ten trials were performed for each dataset
and each $N$, and the mean imbalance is given.  These results are normalized
with respect to $N$, for which the values of $5000$, $50000$, and $500000$ were
chosen.  The `power', `susy', `higgs', and `covertype' datasets are found in the
UCI Machine Learning Repository \cite{ucimlrepository}, the `mnist' dataset is
from \cite{lecun2000mnist}, the `lcdm` and `sdss' datasets are Sloan Digital Sky
Survey data \cite{adelman2008sixth}, and the `randu' dataset is
randomly-generated uniformly-distributed data in 10 dimensions.  The imbalances
on each of these datasets tend to be near-linear.

@@

\begin{table}[tb]
\begin{center}
\begin{tabular}{| c | c | c | c | c |}
\hline
 & & \multicolumn{3}{|c|}{Imbalance} \\
\hline
Dataset & $d$ & $N = 5k$ & $N = 50k$ & $N = 500k$ \\
\hline
%lcdm & 3 & 22417.9 & 257445.6 & 2621239.6 \\
lcdm & 3 & 4.48 & 5.15 & 5.24 \\
% sdss & 4 & 10853.6 & 140431.6 & 1486808.8 \\
sdss & 4 & 2.17 & 2.81 & 2.97 \\
% power & 7 & 27042.2 & 323125.6 & 2251317.2
power & 7 & 5.41 & 6.46 & 4.50 \\
% susy & 18 & 3701.3 & 37823.7 & 431784.8 \\
susy & 18 & 0.74 & 0.76 & 0.86 \\
%randu-10 & 10 & 1156.0 & 10763.4 & 297041.2 \\
randu & 10 & 0.23 & 0.22 & 0.59 \\
% higgs & 29 & 4963.6 & 84171.8 &  \\ need one more...
%covertype & 54 & 6609.6 & 88296.4 & 1247820.7 \\
covertype & 54 & 1.322 & 1.766 & 2.495 \\
% mnist & 784 & 4974.1 & 83658.6 & 1044075.0 \\
mnist & 784 & 0.99 & 1.67 & 2.09 \\
\hline
\end{tabular}
\end{center}
\caption{Empirically calculated tree imbalances, normalized by $N$.}
\label{tab:imbalance}
\end{table}
